### 2021-11-03 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@f43705e294b97501d82ade71305b5b18023fb583 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Food web reconstruction through phylogenetic
transfer of low-rank network representation}
\vskip 5em

%
\href{https://orcid.org/0000-0001-6067-1349}{Tanya\,Strydom}%
%
\,\textsuperscript{1,2,‡}\quad %
\href{https://orcid.org/0000-0003-0193-5441}{Salomé\,Bouskila}%
%
\,\textsuperscript{1,‡}\quad %
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,3,2}\quad %
\href{https://orcid.org/0000-0003-4036-977X}{Ceres\,Barros}%
%
\,\textsuperscript{4}\quad %
\href{https://orcid.org/0000-0002-2151-6693}{Dominique\,Caron}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0003-0452-6993}{Maxwell J\,Farrell}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0002-9935-1366}{Marie-Josée\,Fortin}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0003-3220-6161}{Victoria\,Hemming}%
%
\,\textsuperscript{7}\quad %
\href{https://orcid.org/0000-0002-4104-9463}{Benjamin\,Mercier}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0001-5785-8321}{Rogini\,Runghen}%
%
\,\textsuperscript{8}\quad %
\href{https://orcid.org/0000-0002-3454-0633}{Giulio V.\,Dalla Riva}%
%
\,\textsuperscript{9}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,Département de Sciences Biologiques, Université de
Montréal, Montréal, Canada\quad \textsuperscript{2}\,Quebec Centre for
Biodiversity Science, Montréal,
Canada\quad \textsuperscript{3}\,Département de Biologie, Université de
Sherbrooke, Sherbrooke, Canada\quad \textsuperscript{4}\,Department of
Forest Resources Management, University of British Columbia, Vancouver,
B.C., Canada\quad \textsuperscript{5}\,Department of Biology, McGill
University, Montréal, Canada\quad \textsuperscript{6}\,Department of
Ecology \& Evolutionary Biology, University of Toronto, Toronto,
Canada\quad \textsuperscript{7}\,Department of Forest and Conservation
Sciences, University of British Columbia, Vancouver,
Canada\quad \textsuperscript{8}\,Centre for Integrative Ecology, School
of Biological Sciences, University of Canterbury, Canterbury, New
Zealand\quad \textsuperscript{9}\,School of Mathematics and Statistics,
University of Canterbury, Canterbury, New Zealand

\textsuperscript{‡}\,These authors contributed equally to the work\\

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

\begin{enumerate}
    \itemDespite their importance in many ecological processes,
collecting data and information on ecological interactions is an
exceedingly challenging task. For this reason, large parts of the world
have a data deficit when it comes to species interactions, and how the
resulting networks are structured. As data collection alone is unlikely
to be sufficient, community ecologists must adopt predictive methods.%
    \itemWe present a methodological framework that uses graph embedding
and transfer learning to assemble a predicted list of trophic
interactions of a species pool for which their interactions are unknown.
Specifically, we `learn' the information from a known interaction
network by inferring the latent traits of species and infer the latent
traits of a species pool for which we have no \emph{a priori}
interaction data based on their phylogenetic relatedness to species from
the known network. The latent traits can then be used to predict
interactions and construct and interaction network.%
    \itemHere we assembled a metaweb for Canadian mammals derived from
interactions in the European food web, despite only 4\% of common
species being shared between the two locations. The results of the
predictive model are compared against databases of recorded pairwise
interactions, showing that we correctly recover 91\% of known
interactions.%
    \itemThe framework itself is robust even when the known network is
incomplete or contains spurious interactions making it an ideal
candidate as a tool for filling gaps when it comes to species
interactions. We provide guidance on how this frameworl can be adapted
by substituting some approaches or predictors in order to make it more
generally applicable.%
\end{enumerate}

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen,
2016). First, ecological networks within a location are difficult to
sample properly (Jordano, 2016a, 2016b), resulting in a widespread
``Eltonian shortfall'' (Hortal et al., 2015), \emph{i.e.} a lack of
knowledge about inter- and intra- specific relationships. This first
challenge has been, in large part, addressed by the recent emergence of
a suite of methods aiming to predict interactions within \emph{existing}
networks, many of which are reviewed in Strydom, Catchen, et al. (2021).
Second, recent analyses based on collected data (Poisot, Bergeron, et
al., 2021) or metadata (Cameron et al., 2019) highlight that ecological
networks are currently studied in a biased subset of space and
bioclimates, which impedes our ability to generalize any local
understanding of network structure. Meaning that, although the framework
to address incompleteness \emph{within} networks exists, there would
still be regions for which, due to a \emph{lack} of local interaction
data, we are unable to infer potential species interactions. Having a
general solution for inferring \emph{potential} interactions (despite
the unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales. In a recent overview of
the field of ecological network prediction, Strydom, Catchen, et al.
(2021) identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of
relevant data in most places globally -- paradoxically, this restricts
our ability to infer interactions to locations where inference is
perhaps the least required; second, accurate predictions often demand
accurate predictors, and the lack of methods that can leverage small
amount of data is a serious impediment to our predictive ability
globally.

Here, we present a general method to recommend potential trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang, 2010;
Torrey \& Shavlik, 2010). In this instance, we solve the problem of
predicting trophic interactions between species, based on knowledge
extracted from another species pool for which interactions are known by
using phylogenetic structure as a medium for transfer. There is a
plurality of measures of species similarities that can be used for
inferring \emph{potential} species interactions \emph{i.e.} metaweb
reconstruction (see \emph{e.g.} Morales-Castilla et al., 2015); however,
phylogenetic proximity has several desirable properties when working at
large scales. Gerhold et al. (2015) made the point that phylogenetic
signal captures diversification of characters (large macro-evolutionary
process), but not necessarily community assembly (fine ecological
process); Dormann et al. (2010) previously found very similar
conclusions. Interactions tend to reflect a phylogenetic signal because
they have a conserved pattern of evolutionary convergence that
encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares et al., 2009; Mouquet et al., 2012), and - most
importantly - retain this signal even when it is not detectable at the
community scale (Hutchinson et al., 2017; Poisot \& Stouffer, 2018).
Finally, species interactions at macro-ecological scales seem to respond
mostly to macro-evolutionary processes (Price, 2003); which is evidenced
by the presence of conserved backbones in food webs (Dalla Riva \&
Stouffer, 2016; Mora et al., 2018), strong evolutionary signature on
prey choice (Stouffer et al., 2012), and strong phylogenetic signature
in food web intervality (Eklöf \& Stouffer, 2016). Phylogenetic
reconstruction has also previously been used within the context of
ecological networks, namely understanding ancestral plant-insect
interactions (Braga et al., 2021). Taken together, these considerations
suggest that phylogenies can reliably be used to transfer knowledge on
species interactions.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (latent vulnerability traits are more
representative of species at the lower trophic-level and latent
generality traits are more representative of species at higher
trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (known interactions; Maiorano et al.,
2020a, 2020b) and, based on phylogenetic relationships between mammals
globally (\emph{i.e.}, phylogenetic tree Upham et al., 2019), infer a
metaweb for the Canadian mammalian species pool (interactions are
treated as unknown in this instance). Following the definition of Dunne
(2006), a metaweb is a network analogue to the regional species pool;
specifically, it is an inventory of all \emph{potential} interactions
between species from a spatially delimited area (and so captures the
\(\gamma\) diversity of interactions). The metaweb is, therefore,
\emph{not} a prediction of the food web at a specific locale within the
spatial area it covers, and will have a different structure (notably by
having a larger connectance; see \emph{e.g.} Wood et al., 2015). These
local food webs (which captures the \(\alpha\) diversity of
interactions) are a subset of the metaweb's species and interactions,
and have been called ``metaweb realizations'' (Poisot et al., 2015).
Differences between local food web and their metaweb are due to chance,
species abundance and co-occurrence, local environmental conditions, and
local distribution of functional traits, among others.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it is the
``upper bounds'' on what the composition of the local networks can be
(see \emph{e.g.} McLeod et al., 2021). These local networks, in turn,
can be reconstructed given appropriate knowledge of local species
composition, providing information on structure of food webs at finer
spatial scales. This has been done for example for
tree-galler-parasitoid systems (Gravel et al., 2018), fish trophic
interactions (Albouy et al., 2019), tetrapod trophic interactions
(O'Connor et al., 2020), and crop-pest networks (Grünig et al., 2020).
Whereas the original metaweb definition, and indeed most past uses of
metawebs, was based on the presence/absence of interactions, we focus on
\emph{probabilistic} metawebs where interactions are represented as the
chance of success of a Bernoulli trial (see \emph{e.g.} Poisot et al.,
2016); therefore, not only does our method recommend interactions that
may exist, it gives each interaction a score, allowing us to properly
weigh them.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%. It should be reiterated that the framework presented
in fig.~\ref{fig:concept} is amenable to changes; notably, the measure
of similarity may not be phylogeny, and can be replaced by information
on foraging (Beckerman et al., 2006), cell-level mechanisms (Boeckaerts
et al., 2021), or a combination of traits and phylogenetic structure
(Stock, 2021). Most importantly, although we focus on a trophic system,
it is an established fact that different (non-trophic) interactions do
themselves interact with and influence the outcome of trophic
interactions (see \emph{e.g.} Kawatsu et al., 2021; Kéfi et al., 2012).
Future development of metaweb inference techniques should cover the
prediction of multiple interaction types.

\hypertarget{data-used-for-the-case-study}{%
\section{Data used for the case
study}\label{data-used-for-the-case-study}}

We use data from the European metaweb assembled by Maiorano et al.
(2020a). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
This metaweb itself is a network of binary (\emph{i.e.}
presence/absence), potential two-way interactions between species pairs.

We filtered down the European metaweb to create a subgraph corresponding
to all mammals by matching species names in the original network to the
Global Biodiversity Information Facility (GBIF) taxonomic backbone (GBIF
Secretariat, 2021) and retaining all those who matched to mammals. This
serves a dual purpose 1) to extract only mammals from the European
network and 2) to match and standardize species names when aggregating
the different data sources further downstream (which is an important
consideration when combining datasets (Grenié et al., 2021)). All nodes
had valid matches to GBIF at this step, and so this backbone is used for
all name reconciliation steps as outlined below.

The European metaweb represents the knowledge we want to learn and
transfer; the phylogenetic similarity of mammals here represents the
information for transfer (\emph{i.e.} the transfer medium). We used the
mammalian consensus supertree by Upham et al. (2019), for which all
approximatively 6000 names have been similarly matched to their GBIF
valid names. This step allows us to place each node of the mammalian
European metaweb in the phylogeny.

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the International Union for Conservation of Nature
(IUCN) checklist, and selected the terrestrial and semi-aquatic species
(this corresponds to the same selection that was applied by Maiorano et
al. (2020a) in the European metaweb). The IUCN names were, as
previously, reconciled against GBIF to have an exact match to the
taxonomy.

After taxonomic cleaning and reconciliation as outlined in the following
sections, the mammalian European metaweb has 260 species, and the
Canadian species pool has 163; of these, 17 (about 4\% of the total) are
shared, and 89 species from Canada (54\%) had at least one congeneric
species in Europe. The similarity for both species pools predictably
increases with higher taxonomic order, with 19\% of shared genera, 47\%
of shared families, and 75\% of shared orders; for the last point,
Canada and Europe each had a single unique order (\emph{Didelphimorphia}
for Canada, \emph{Erinaceomorpha} for Europe).

In the following sections, we describe the representational learning
step applied to European data, the transfer step through phylogenetic
similarity, and the generation of a probabilistic metaweb for the
destination species pool.

\hypertarget{method-description}{%
\section{Method description}\label{method-description}}

The core point of our method is the transfer of knowledge of a known
ecological network, in order to predict interactions between species
from another location at which the network is unknown (or partially
known). In fig.~\ref{fig:concept}, we give a high-level overview of the
approach; in the example around which this manuscript is built
(leveraging detailed knowledge about binary trophic interactions between
Mammalia in Europe to predict the less known trophic interactions
between closely phylogenetically related Mammalia in Canada), we use a
series of specific steps for network embedding, trait inference, network
prediction and thresholding.

Specifically, our approach can be summarized as follows: from the known
network in Europe, we use a truncated Singular Value Decomposition
(t-SVD; Halko et al., 2011) to generate latent traits representing a
low-dimensional embedding of the network. As an aside, most ecologists
are indirectly familiar with SVD: Principal Component Analysis is a
special case of SVD, which is more sensitive to numerical instabilities
(see notably Shlens, 2014). The latent traits give an unbiased estimate
of the node's position in the latent feature spaces and can be mapped
onto a reference phylogeny (other distance-based measures of species
proximity that allow for the inference of features in the latent space
can be used, for example the dissimilarity in functional traits). Based
on the reconstructed latent traits for species in the destination
species pool, a Random Dot Product Graph model (hereafter RDPG; S. J.
Young \& Scheinerman, 2007) predicts the interaction between species
through a function of the nodes' features through matrix multiplication.
Thus, from latent traits and node position, we can infer interactions.

The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
(Holm, 2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information than we might think; as such, even algorithms that are
disconnected from the model can make correct guesses most of the time
(Halevy et al., 2009); in fact, in an instance of ecological forecasting
of spatio-temporal systems, model-free approaches (\emph{i.e.} drawing
all of their information from the data) outperformed model-informed ones
(Perretti et al., 2013).

\hypertarget{implementation-and-code-availability}{%
\subsection{Implementation and code
availability}\label{implementation-and-code-availability}}

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson et al.,
2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl} (Dansereau \&
Poisot, 2021). The network embedding and analysis is done using
\texttt{EcologicalNetworks.jl} (Banville et al., 2021; Poisot et al.,
2019). The phylogenetic simulations are done using
\texttt{PhyloNetworks.jl} (Solís-Lemus et al., 2017) and
\texttt{Phylo.jl} (Reeve et al., 2016). A complete \texttt{Project.toml}
file specifying the full tree of dependencies is available alongside the
code. This material also includes a fully annotated copy of the entire
code required to run this project (describing both the intent of the
code and discussing some technical implementation details), a vignette
for every step of the process, and a series of Jupyter notebooks with
the text and code. The pipeline can be executed on a laptop in a matter
of minutes, and therefore does not require extensive computational
power.

\hypertarget{step-1-learning-the-origin-network-representation}{%
\subsection{Step 1: Learning the origin network
representation}\label{step-1-learning-the-origin-network-representation}}

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy et al., 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer, 2016); in other words, the latent variables of an RDPG
can be mapped onto a phylogenetic tree, and phylogenetically similar
predators should share phylogenetically similar preys. In addition,
recent advances show that the latent variables produced this way can be
used to predict \emph{de novo} network edges. Interestingly, the latent
variables do not need to be produced by decomposing the network itself;
in a recent contribution, Runghen et al. (2021) showed that deep
artificial neural networks are able to reconstruct the left and right
subspaces of an RDPG, in order to predict human movement networks from
individual/location metadata. This is an exciting opportunity, as it
opens up the possibility of using additional metadata as predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom, Dalla
Riva, et al., 2021) and to allow highly accurate prediction of the
interactions within a single network (Poisot, Ouellet, et al., 2021).
Under SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.

The specific rank at which the SVD ought to be truncated is a difficult
question. The purpose of SVD is to remove the noise (expressed at high
dimensions) and to focus on the signal, (expressed at low dimensions).
In datasets with a clear signal/noise demarcation, a scree plot of
\(\mathbf{\Sigma}\) can show a sharp drop at the rank where noise starts
(Zhu \& Ghodsi, 2006). Because the European metaweb is almost entirely
known, the amount of noise (uncertainty) is low; this is reflected in
fig.~\ref{fig:scree} (left), where the scree plot shows no important
drop, and in fig.~\ref{fig:scree} (right) where the proportion of
variance explained increases smoothly at higher dimensions. For this
reason, we default back to a threshold that explains 60\% of the
variance in the underlying data, corresponding to 12 dimensions -
\emph{i.e.} a tradeoff between accuracy and a reduced number of
features.

An RDPG estimates the probability of observing interactions between
nodes (species) as a function of the nodes' latent variables, and is a
way to turn a SVD (which decompose one matrix into three) into two
matrices that can be multiplied to provide an approximation of the
network. The latent variables used for the RDPG, called the left and
right subspaces, are defined as
\(\mathscr{L} = \mathbf{U}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{V}'\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R} = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R} \approx \mathbf{A}\). Using a rank of 1 for the
t-SVD provides a first-order approximation of the network. One advantage
of using a RDPG rather than a SVD is that the number of components to
estimate decreases; notably, one does not have to estimate the singular
values of the SVD. Furthermore, the two subspaces can be directly
multiplied to yield a network.

\begin{figure}
\hypertarget{fig:scree}{%
\centering
\includegraphics{figures/figure-screeplot.png}
\caption{Left: representation of the scree plot of the singular values
from the t-SVD on the European metaweb. The scree plot shows no obvious
drop in the singular values that may be leveraged to automatically
detect a minimal dimension for embedding, after \emph{e.g.} Zhu \&
Ghodsi (2006). Right: cumulative fraction of variance explained by each
dimension up to the rank of the European metaweb. The grey lines
represent cutoffs at 50, 60, \ldots, 90\% of variance explained. For the
rest of the analysis, we reverted to an arbitrary threshold of 60\% of
variance explained, which represented a good tradeoff between accuracy
and reduced number of features.}\label{fig:scree}
}
\end{figure}

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
Maiorano et al. (2020a) and O'Connor et al. (2020), this seems like a
reasonable assumption, although one that does not hold for all metawebs.
We used the thresholding approach presented in Poisot, Ouellet, et al.
(2021), and picked a cutoff that maximized Youden's \(J\) statistic (a
measure of the informedness (trust) of predictions; Youden (1950)); the
resulting cutoff was 0.22, and gave an accuracy above 0.99. In Supp.
Mat. 1, we provide several lines of evidence that using the entire
network to estimate the threshold does not lead to overfitting; that
using a subset of species would yield the same threshold; that
decreasing the quality of the original data by adding of removing
interactions would minimally affect the predictive accuracy of RDPG
applied to the European metaweb; and that the networks reconstructed
from artificially modified data are reconstructed with the correct
ecological properties.

The left and right subspaces for the European metaweb, accompanied by
the threshold for prediction, represent the knowledge we seek to
transfer. In the next section, we explain how we rely on phylogenetic
similarity to do so.

\hypertarget{steps-2-and-3-transfer-learning-through-phylogenetic-relatedness}{%
\subsection{Steps 2 and 3: Transfer learning through phylogenetic
relatedness}\label{steps-2-and-3-transfer-learning-through-phylogenetic-relatedness}}

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin, 2012). This uses the
estimated feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using latent generality and
vulnerability traits) and allows us to impute the missing (latent) trait
data for the Canadian species that are not already in the European
network; as we are focused on predicting contemporary interactions, we
only retained the values for the tips of the tree. We assumed that all
traits (\emph{i.e.} the feature vectors for the left and right
subspaces) were independent, which is a reasonable assumption as every
trait/dimension added to the t-SVD has an \emph{additive} effect to the
one before it. Note that the Upham et al. (2019) tree itself has some
uncertainty associated to inner nodes of the phylogeny. In this case
study, we have decided to not propagate this uncertainty, as it would
complexify the process. The Brownian motion algorithm returns the
\emph{average} value of the trait, and its upper and lower bounds.
Because we do not estimate other parameters of the traits'
distributions, we considered that every species trait is represented as
a uniform distribution between these bounds. The choice of the uniform
distribution was made because the algorithm returns a minimum and
maximum point estimate for the value, and given this information, the
uniform distribution is the one with maximum entropy. Had all mean
parameters estimates been positive, the exponential distribution would
have been an alternative, but this is not the case for the subspaces of
an RDPG. In order to examine the consequences of the choice of
distribution, we estimated the variance per latent variable per node to
use a Normal distribution; as we show in Supp. Mat. 2, this decision
results in dramatically over-estimating the number and probability of
interactions, and therefore we keep the discussions in the main text to
the uniform case. The inferred left and right subspaces for the Canadian
species pool (\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\)) have
entries that are distributions, representing the range of values for a
given species at a given dimension.

These objects represent the transferred knowledge, which we can use for
prediction of the Canadian metaweb.

\hypertarget{step-4-probabilistic-prediction-of-the-destination-network}{%
\subsection{Step 4: Probabilistic prediction of the destination
network}\label{step-4-probabilistic-prediction-of-the-destination-network}}

The phylogenetic reconstruction of \(\hat{\mathscr{L}}\) and
\(\hat{\mathscr{R}}\) has an associated uncertainty, represented by the
breadth of the uniform distribution associated to each of their entries.
Therefore, we can use this information to assemble a
\emph{probabilistic} metaweb in the sense of Poisot et al. (2016),
\emph{i.e.} in which every interaction is represented as a single,
independent, Bernoulli event of probability \(p\).

\begin{figure}
\hypertarget{fig:subspaces}{%
\centering
\includegraphics{figures/figure-subspaces.png}
\caption{Visual representation of the left (green/purple) and right
(green/brown) subspaces, alongside the adjacency matrix of the food web
they encode (greyscale). The European metaweb is on the left, and the
imputed Canadian metaweb (before data inflation) on the right. This
figure illustrates how much structure the left subspace captures. As we
show in fig.~\ref{fig:degree}, the species with a value of 0 in the left
subspace are species without any prey.}\label{fig:subspaces}
}
\end{figure}

Specifically, we have adopted the following approach. For every entry in
\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\), we draw a value from
its distribution. This results in one instance of the possible left
(\(\hat{\mathscr{l}}\)) and right (\(\hat{\mathscr{r}}\)) subspaces for
the Canadian metaweb. These can be multiplied, to produce one matrix of
real values. Because the entries in \(\hat{\mathscr{l}}\) and
\(\hat{\mathscr{r}}\) are in the same space where \(\mathscr{L}\) and
\(\mathscr{R}\) were originally predicted, it follows that the threshold
\(\rho\) estimated for the European metaweb also applies. We use this
information to produce one random Canadian metaweb,
\(N = \hat{\mathscr{L}}\hat{\mathscr{R}}' \ge \rho\). As we can see in
(fig.~\ref{fig:subspaces}), the European and Canadian metawebs are
structurally similar (as would be expected given the biogeographic
similarities) and the two (left and right) subspaces are distinct
\emph{i.e.} capturing predation (generality) and prey (vulnerability)
latent traits.

Because the intervals around some trait values can be broad (in fact,
probably broader than what they would actually be, see \emph{e.g.}
Garland et al., 1999), we repeat the above process \(2\times 10^5\)
times, which results in a probabilistic metaweb \(P\), where the
probability of an interaction (here conveying our degree of trust that
it exists given the inferred trait distributions) is given by the number
of times where it appears across all random draws \(N\), divided by the
number of samples. An interaction with \(P_{i,j} = 1\) means that these
two species were predicted to interact in all \(2\times 10^5\) random
draws.

It must be noted that despite bringing in a large amount of information
from the European species pool and interactions, the Canadian metaweb
has distinct structural properties. Following an approach similar to
Vermaat et al. (2009), we show in Supp. Mat. 3 that not only can we
observe differences in a multivariate space between the European and
Canadian metaweb, we can also observe differences in the same space
between random subgraphs from these networks. These results line up with
the studies spatializing metawebs that have been discussed in the
introduction: changes in the species pool are driving local structural
changes in the networks.

\hypertarget{data-cleanup-discovery-validation-and-thresholding}{%
\subsection{Data cleanup, discovery, validation, and
thresholding}\label{data-cleanup-discovery-validation-and-thresholding}}

Once the probabilistic metaweb for Canada has been produced, we followed
a number of data inflation steps to finalize it. This step is external
to the actual transfer learning framework but rather serves as a way to
augment and validate the predicted metaweb.

\begin{figure}
\hypertarget{fig:inflation}{%
\centering
\includegraphics{figures/figure-validation.png}
\caption{Left, comparison of the probabilities of interactions assigned
by the model to all interactions (grey curve), the subset of
interactions found in GLOBI (red), and in the Strong \& Leroux (2014)
Newfoundland dataset (blue). The model recovers more interactions with a
low probability compared to data mining, which can suggest that
collected datasets are biased towards more common or easy to identify
interactions. Right, distribution of the in-degree and out-degree of the
mammals from Canada in the reconstructed metaweb. This figure describes
a flat, relatively short food web, in which there are few predators but
a large number of preys.}\label{fig:inflation}
}
\end{figure}

First, we extracted the subgraph corresponding to the 17 species shared
between the European and Canadian pools and replaced these interactions
with a probability of 0 (non-interaction) or 1 (interaction), according
to their value in the European metaweb. This represents a minute
modification of the inferred network (about 0.8\% of all species pairs
from the Canadian web), but ensures that we are directly re-using
knowledge from Europe.

Second, we looked for all species in the Canadian pool known to the
Global Biotic Interactions (GLoBI) database (Poelen et al., 2014), and
extracted their known interactions. Because GLoBI aggregates observed
interactions, it is not a \emph{networks} data source, and therefore the
only information we can reliably extract from it is that a species pair
\emph{was reported to interact at least once}. This last statement
should yet be taken with caution, as some sources in GLoBI (\emph{e.g.}
Thessen \& Parr, 2014) are produced through text analysis, and therefore
may not document direct evidence of the interaction. Nevertheless,
should the predictive model work, we would expect that a majority of
interactions known to GLoBI would also be predicted. We retrieved 366
interactions between mammals from the Canadian species pool from GLoBI,
33 of which were not predicted by the model; this results in a success
rate of 91\%. After performing this check, we set the probability of all
interactions known to GLoBI to 1.

Finally, we downloaded the data from Strong \& Leroux (2014), who mined
various literature sources to identify trophic interactions in
Newfoundland. This dataset documented 25 interactions between mammals,
only two of which were not part of our (Canada-level) predictions,
resulting in a success rate of 92\%. These two interactions were added
to our predicted metaweb with a probability of 1. A table listing all
interactions in the predicted Canadian metaweb can be found in the
supplementary material.

\begin{figure}
\hypertarget{fig:thresholds}{%
\centering
\includegraphics{figures/figure-cutoffs.png}
\caption{Left: effect of varying the cutoff for probabilities to be
considered non-zero on the number of unique links and on \(\hat{L}\),
the probabilistic estimate of the number of links assuming that all
interactions are independent. Right: effect of varying the cutoff on the
number of disconnected species, and on network connectance. In both
panels, the grey line indicates the cutoff
\(P(i\rightarrow j) \approx 0.08\) that resulted in the first species
losing all of its interactions.}\label{fig:thresholds}
}
\end{figure}

Because the confidence intervals on the inferred trait space are
probably over-estimates, we decided to apply a thresholding step to the
interactions after the data inflation (fig.~\ref{fig:thresholds}).
Cirtwill \& Hambäck (2021) proposed a number of strategies to threshold
probabilistic networks. Their methods assume the underlying data to be
tag-based sequencing, which represents interactions as co-occurrences of
predator and prey within the same tags; this is conceptually identical
to our Bernoulli-trial based reconstruction of a probabilistic network.
We performed a full analysis of the effect of various cutoffs, and as
they either resulted in removing too few interactions, or removing
enough interactions that species started to be disconnected from the
network, we set this threshold for a probability equivalent to 0 to the
largest possible value that still allowed all species to have at least
one interaction with a non-zero probability. The need for this slight
deviation from the Cirtwill \& Hambäck (2021) method highlights the need
for additional development on network thresholding.

\hypertarget{results-and-discussion-of-the-case-study}{%
\section{Results and discussion of the case
study}\label{results-and-discussion-of-the-case-study}}

In fig.~\ref{fig:thresholds}, we examine the effect of varying the
cutoff on \(P(i \rightarrow j)\) on the number of links, species, and
connectance. Determining a cutoff using the maximum curvature, or
central difference approximation of the second order partial derivative,
as suggested by \emph{e.g.} Cirtwill \& Hambäck (2021), results in
species being lost, or almost all links being kept. We therefore settled
on the value that allowed all species to remain with at least one
interaction. This result, in and of itself, suggests that additional
methodological developments for the thresholding of probabilistic
networks are required.

\begin{figure}
\hypertarget{fig:degree}{%
\centering
\includegraphics{figures/figure-degree.png}
\caption{Top: biological significance of the first dimension. Left:
there is a linear relationship between the values on the first dimension
of the left subspace and the generality, \emph{i.e.} the relative number
of preys, \emph{sensu} Schoener (1989). Species with a value of 0 in
this subspace are at the bottom-most trophic level. Right: there is,
similarly, a linear relationship between the position of a species on
the first dimension of the right subspace and its vulnerability,
\emph{i.e.} the relative number of predators. Taken together, these two
figures show that the first-order representation of this network would
capture its degree distribution. Bottom: topological consequences of the
first dimension. Left: differences in the \(z\)-score of the actual
configuration model for the reconstructed network, and the prediction
based only on the first dimension. Right: distribution of the
differences in the left panel.}\label{fig:degree}
}
\end{figure}

The t-SVD embedding is able to learn relevant ecological features for
the network. fig.~\ref{fig:degree} shows that the first rank correlates
linearly with generality and vulnerability (Schoener, 1989), \emph{i.e.}
the number of preys and predators for each species. Importantly, this
implies that a rank 1 approximation represents the configuration model
for the metaweb, \emph{i.e.} a set of random networks generated from a
given degree sequence (Park \& Newman, 2004). Accounting for the
probabilistic nature of the degrees, the rank 1 approximation also
represents the \emph{soft} configuration model (van der Hoorn et al.,
2018). Both models are maximum entropy graph models (Garlaschelli et
al., 2018), with sharp (all network realizations satisfy the specified
degree sequence) and soft (network realizations satisfy the degree
sequence on average) local constraints, respectively. The (soft)
configuration model is an unbiased random graph model widely used by
ecologists in the context of null hypothesis significance testing of
network structure (\emph{e.g.} Bascompte et al., 2003) and can provide
informative priors for Bayesian inference of network structure
(\emph{e.g.} J.-G. Young et al., 2021). It is noteworthy that for this
metaweb, the relevant information was extracted at the first rank.
Because the first rank corresponds to the leading singular value of the
system, the results of fig.~\ref{fig:degree} have a straightforward
interpretation: degree-based processes are the most important in
structuring the mammalian food web.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

One important aspect in which Europe and Canada differ (despite their
comparable bioclimatic conditions) is the degree of the legacy of human
impacts, which have been much longer in Europe. Nenzén et al. (2014)
showed that even at small scales (the Iberian peninsula), mammal food
webs retain the signal of both climate change and human activity, even
when this human activity was orders of magnitude less important than it
is now. Similarly, Yeakel et al. (2014) showed that changes in human
occupation over several centuries can lead to food web collapse.
Megafauna in particular seems to be very sensitive to human arrival
(Pires et al., 2015). In short, there is well-substantiated support for
the idea that human footprint affects more than the risk of species
extinction (Marco et al., 2018), and can lead to changes in interaction
structure. Yet, owing to the inherent plasticity of interactions, there
have been documented instances of food webs undergoing rapid
collapse/recovery cycles over short periods of time (Pedersen et al.,
2017). The embedding of a network, in a sense, embeds its
macro-evolutionary history, especially as RDPG captures ecological
signal (Dalla Riva \& Stouffer, 2016); at this point, it is important to
recall that a metaweb is intended as a catalogue of all potential
interactions, which should then be filtered (Morales-Castilla et al.,
2015). In practice (and in this instance) the reconstructed metaweb will
predict interactions that are plausible based on the species'
evolutionary history, however some interactions would/would not be
realized due to human impact.

Dallas et al. (2017) suggested that most links in ecological networks
may be cryptic, \emph{i.e.} uncommon or otherwise hard to observe. This
argument essentially echoes Jordano (2016b): the sampling of ecological
interactions is difficult because it requires first the joint
observation of two species, and then the observation of their
interaction. In addition, it is generally expected that weak or rare
links would be more common in networks (Csermely, 2004), compared to
strong, persistent links; this is notably the case in food chains,
wherein many weaker links are key to the stability of a system (Neutel
et al., 2002). In the light of these observations, the results in
fig.~\ref{fig:inflation} are not particularly surprising: we expect to
see a surge in these low-probability interactions under a model that has
a good predictive accuracy. Because the predictions we generate are by
design probabilistic, then one can weigh these rare links appropriately.
In a sense, that most ecological interactions are elusive can call for a
slightly different approach to sampling: once the common interactions
are documented, the effort required in documenting each rare interaction
may increase exponentially. Recent proposals suggest that machine
learning algorithms, in these situations, can act as data generators
(Hoffmann et al., 2019): in this perspective, high quality observational
data can be supplemented with synthetic data coming from predictive
models, which increases the volume of information available for
inference. Indeed, Strydom, Catchen, et al. (2021) suggested that
knowing the metaweb may render the prediction of local networks easier,
because it fixes an ``upper bound'' on which interactions can exist;
indeed, with a probabilistic metaweb, we can consider that the metaweb
represents an aggregation of informative priors on the interactions.

Related to the last point, Cirtwill et al. (2019) showed that network
inference techniques based on Bayesian approaches would perform far
better in the presence of an interaction-level informative prior; the
desirable properties of such a prior would be that it is expressed as a
probability, preferably representing a Bernoulli event, the value of
which would be representative of relevant biological processes
(probability of predation in this case). We argue that the probability
returned at the very last step of our framework may serve as this
informative prior; indeed, the output of our analysis can be used in
subsequent steps, also possibly involving expert elicitation to validate
some of the most strongly recommended interactions. One important
\emph{caveat} to keep in mind when working with interaction inference is
that interactions can never really be true negatives (in the current
state of our methodological framework and data collection limitations);
this renders the task of validating a model through the usual
application of binary classification statistics very difficult (although
see Strydom, Catchen, et al., 2021 for a discussion of alternative
suggestions). The other way through which our framework can be improved
is by substituting the predictors that are used for transfer. For
example, in the presence of information on species traits that are known
to be predictive of species interactions, one might want to rely on
functional rather than phylogenetic distances -- in food webs, body size
(and allometrically related variables) has been established as such a
variable (Brose et al., 2006); the identification of relevant functional
traits is facilitated by recent methodological developments (Rosado et
al., 2013). It should be noted that Xing \& Fayle (2021) highlight
phylogenetic relatedness as one of the core components of network
comparison at the global scale. In this case study, we have embedded the
original metaweb using t-SVD, because it lends itself to an RDPG
reconstruction, which is known to capture the consequences of
evolutionary processes (Dalla Riva \& Stouffer, 2016); this being said,
there are other ways to embed graphs (Arsov \& Mirceva, 2019; Cai et
al., 2017; Cao et al., 2019), which can be used as alternatives.

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, our approach (and
indeed, any inference of a metaweb at large scales) must contend with
several interesting and interwoven families of problems. The first is
the limit of the metaweb to embed and transfer. If the initial metaweb
is too narrow in scope, notably from a taxonomic point of view, the
chances of finding another area with enough related species to make a
reliable inference decreases; this would likely be indicated by large
confidence intervals during ancestral character estimation, but the lack
of well documented metawebs is currently preventing the development of
more concrete guidelines. The question of phylogenetic relatedness and
dispersal is notably true if the metaweb is assembled in an area with
mostly endemic species, and as with every predictive algorithm, there is
room for the application of our best ecological judgement. Conversely,
the metaweb should be reliably filled, which assumes that the \(S^2\)
interactions in a pool of \(S\) species have been examined, either
through literature surveys or expert elicitation. Supp. Mat. 1 provides
some guidance as to the type of sampling effort that should be
prioritized. Although RDPG was able to maintain very high predictive
power when interactions were missing, the addition of false positive
interactions was immediately detected; this suggests that it may be
appropriate to err on the side of ``too many'' interactions when
constructing the initial metaweb to be transferred. The second series of
problems are related to determining which area should be used to infer
the new metaweb in, as this determines the species pool that must be
used. In our application, we focused on the mammals of Canada. The
upside of this approach is that information at the country level is
likely to be required by policy makers and stakeholders for their
biodiversity assessment, as each country tends to set goals at the
national level (Buxton et al., 2021) for which quantitative instruments
are designed (Turak et al., 2017), with specific strategies often
enacted at smaller scales (Ray et al., 2021). And yet, we do not really
have a satisfying answer to the question of ``where does a food web
stop?''; the current most satisfying solutions involve examining the
spatial consistency of network area relationships (Fortin et al., 2021;
see \emph{e.g.} Galiana et al., 2018, 2019, 2021), which is of course
impossible in the absence of enough information about the network
itself. This suggests that an \emph{a posteriori} refinement of the
results may be required, based on a downscaling of the metaweb. The
final family of problems relates less to the availability of data or
quantitative tools, and more to the praxis of spatial ecology. Operating
under the context of national divisions, in large parts of the world,
reflects nothing more than the legacy of settler colonialism. Indeed,
the use of ecological data is not an apolitical act (Nost \& Goldstein,
2021), as data infrastructures tend to be designed to answer questions
within national boundaries, and their use both draws upon and reinforces
territorial statecraft; as per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that methods such as we
propose operate under the framework that contributed to the ongoing
biodiversity crisis (Adam, 2014), reinforced environmental injustice
(Choudry, 2013; Domínguez \& Luoma, 2020), and on Turtle Island
especially, should be replaced by Indigenous principles of land
management (Eichhorn et al., 2019; No'kmaq et al., 2021). As we see
AI/ML being increasingly mobilized to generate knowledge that is lacking
for conservation decisions (\emph{e.g.} Lamba et al., 2019; Mosebo
Fernandes et al., 2020), our discussion of these tools need to go beyond
the technical, and into the governance consequences they can have.

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. TP, TS, DC, and LP received funding from the Canadian Institue
for Ecology \& Evolution. FB is funded by the Institute for Data
Valorization (IVADO). TS, SB, and TP are funded by a donation from the
Courtois Foundation. CB was awarded a Mitacs Elevate Fellowship no.
IT12391, in partnership with fRI Research, and also acknowledges funding
from Alberta Innovates and the Forest Resources Improvement Association
of Alberta. M-JF acknowledges funding from NSERC Discovery Grant and
NSERC CRC. RR is funded by New Zealand's Biological Heritage Ngā Koiora
Tuku Iho National Science Challenge, administered by New Zealand
Ministry of Business, Innovation, and Employment. BM is funded by the
NSERC Alexander Graham Bell Canada Graduate Scholarship and the FRQNT
master's scholarship. LP acknowledges funding from NSERC Discovery Grant
(NSERC RGPIN-2019-05771). TP acknowledges financial support from NSERC
through the Discovery Grants and Discovery Accelerator Supplement
programs.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Adam2014EleTre}{}%
Adam, R. (2014). \emph{Elephant treaties: The Colonial legacy of the
biodiversity crisis}. UPNE.

\leavevmode\hypertarget{ref-Albouy2019MarFis}{}%
Albouy, C., Archambault, P., Appeltans, W., Araújo, M. B., Beauchesne,
D., Cazelles, K., Cirtwill, A. R., Fortin, M.-J., Galiana, N., Leroux,
S. J., Pellissier, L., Poisot, T., Stouffer, D. B., Wood, S. A., \&
Gravel, D. (2019). The marine fish food web is globally connected.
\emph{Nature Ecology \& Evolution}, \emph{3}(8, 8), 1153--1161.
\url{https://doi.org/10.1038/s41559-019-0950-y}

\leavevmode\hypertarget{ref-Arsov2019NetEmb}{}%
Arsov, N., \& Mirceva, G. (2019, November 26). \emph{Network Embedding:
An Overview}. \url{http://arxiv.org/abs/1911.11726}

\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S., \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software},
\emph{6}(61), 2721. \url{https://doi.org/10.21105/joss.02721}

\leavevmode\hypertarget{ref-Bascompte2003NesAss}{}%
Bascompte, J., Jordano, P., Melian, C. J., \& Olesen, J. M. (2003). The
nested assembly of plant-animal mutualistic networks. \emph{Proceedings
of the National Academy of Sciences}, \emph{100}(16), 9383--9387.
\url{https://doi.org/10.1073/pnas.1633576100}

\leavevmode\hypertarget{ref-Beckerman2006ForBio}{}%
Beckerman, A. P., Petchey, O. L., \& Warren, P. H. (2006). Foraging
biology predicts food web complexity. \emph{Proceedings of the National
Academy of Sciences}, \emph{103}(37), 13745--13749.
\url{https://doi.org/10.1073/pnas.0603039103}

\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
Fresh Approach to Numerical Computing. \emph{SIAM Review}, \emph{59}(1),
65--98. \url{https://doi.org/10.1137/141000671}

\leavevmode\hypertarget{ref-Boeckaerts2021PreBac}{}%
Boeckaerts, D., Stock, M., Criel, B., Gerstmans, H., De Baets, B., \&
Briers, Y. (2021). Predicting bacteriophage hosts based on sequences of
annotated receptor-binding proteins. \emph{Scientific Reports},
\emph{11}(1, 1), 1467. \url{https://doi.org/10.1038/s41598-021-81063-4}

\leavevmode\hypertarget{ref-Braga2021PhyRec}{}%
Braga, M. P., Janz, N., Nylin, S., Ronquist, F., \& Landis, M. J.
(2021). Phylogenetic reconstruction of ancestral ecological networks
through time for pierid butterflies and their host plants. \emph{Ecology
Letters}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/ele.13842}

\leavevmode\hypertarget{ref-Brose2006ConRes}{}%
Brose, U., Jonsson, T., Berlow, E. L., Warren, P., Banasek-Richter, C.,
Bersier, L.-F., Blanchard, J. L., Brey, T., Carpenter, S. R.,
Blandenier, M.-F. C., Cushing, L., Dawah, H. A., Dell, T., Edwards, F.,
Harper-Smith, S., Jacob, U., Ledger, M. E., Martinez, N. D., Memmott,
J., \ldots{} Cohen, J. E. (2006). ConsumerResource Body-Size
Relationships in Natural Food Webs. \emph{Ecology}, \emph{87}(10),
2411--2417.
\url{https://doi.org/10.1890/0012-9658(2006)87\%5B2411:CBRINF\%5D2.0.CO;2}

\leavevmode\hypertarget{ref-Buxton2021KeyInf}{}%
Buxton, R. T., Bennett, J. R., Reid, A. J., Shulman, C., Cooke, S. J.,
Francis, C. M., Nyboer, E. A., Pritchard, G., Binley, A. D., Avery-Gomm,
S., Ban, N. C., Beazley, K. F., Bennett, E., Blight, L. K., Bortolotti,
L. E., Camfield, A. F., Gadallah, F., Jacob, A. L., Naujokaitis-Lewis,
I., \ldots{} Smith, P. A. (2021). Key information needs to move from
knowledge to action for biodiversity conservation in Canada.
\emph{Biological Conservation}, \emph{256}, 108983.
\url{https://doi.org/10.1016/j.biocon.2021.108983}

\leavevmode\hypertarget{ref-Cai2017ComSur}{}%
Cai, H., Zheng, V. W., \& Chang, K. C.-C. (2017). \emph{A Comprehensive
Survey of Graph Embedding: Problems, Techniques and Applications}.
\url{http://arxiv.org/abs/1709.07604}

\leavevmode\hypertarget{ref-Cameron2019UneGlo}{}%
Cameron, E. K., Sundqvist, M. K., Keith, S. A., CaraDonna, P. J.,
Mousing, E. A., Nilsson, K. A., Metcalfe, D. B., \& Classen, A. T.
(2019). Uneven global distribution of food web studies under climate
change. \emph{Ecosphere}, \emph{10}(3), e02645.
\url{https://doi.org/10.1002/ecs2.2645}

\leavevmode\hypertarget{ref-Cao2019NetEmb}{}%
Cao, R.-M., Liu, S.-Y., \& Xu, X.-K. (2019). Network embedding for link
prediction: The pitfall and improvement. \emph{Chaos: An
Interdisciplinary Journal of Nonlinear Science}, \emph{29}(10), 103102.
\url{https://doi.org/10.1063/1.5120724}

\leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
Cavender-Bares, J., Kozak, K. H., Fine, P. V. A., \& Kembel, S. W.
(2009). The merging of community ecology and phylogenetic biology.
\emph{Ecology Letters}, \emph{12}(7), 693--715.
\url{https://doi.org/10.1111/j.1461-0248.2009.01314.x}

\leavevmode\hypertarget{ref-Choudry2013SavBio}{}%
Choudry, A. (2013). Saving biodiversity, for whom and for what?
Conservation NGOs, complicity, colonialism and conquest in an era of
capitalist globalization. In \emph{NGOization: Complicity,
contradictions and prospects} (pp. 24--44). Bloomsbury Publishing.

\leavevmode\hypertarget{ref-Cirtwill2019QuaFra}{}%
Cirtwill, A. R., Ekl, A., Roslin, T., Wootton, K., \& Gravel, D. (2019).
A quantitative framework for investigating the reliability of empirical
network construction. \emph{Methods in Ecology and Evolution}, \emph{0}.
\url{https://doi.org/10.1111/2041-210X.13180}

\leavevmode\hypertarget{ref-Cirtwill2021BuiFoo}{}%
Cirtwill, A. R., \& Hambäck, P. (2021). Building food networks from
molecular data: Bayesian or fixed-number thresholds for including links.
\emph{Basic and Applied Ecology}, \emph{50}, 67--76.
\url{https://doi.org/10.1016/j.baae.2020.11.007}

\leavevmode\hypertarget{ref-Csermely2004StrLin}{}%
Csermely, P. (2004). Strong links are important, but weak links
stabilize them. \emph{Trends in Biochemical Sciences}, \emph{29}(7),
331--334. \url{https://doi.org/10.1016/j.tibs.2004.05.004}

\leavevmode\hypertarget{ref-DallaRiva2016ExpEvo}{}%
Dalla Riva, G. V., \& Stouffer, D. B. (2016). Exploring the evolutionary
signature of food webs' backbones using functional traits. \emph{Oikos},
\emph{125}(4), 446--456. \url{https://doi.org/10.1111/oik.02305}

\leavevmode\hypertarget{ref-Dallas2017PreCry}{}%
Dallas, T., Park, A. W., \& Drake, J. M. (2017). Predicting cryptic
links in host-parasite networks. \emph{PLOS Computational Biology},
\emph{13}(5), e1005557.
\url{https://doi.org/10.1371/journal.pcbi.1005557}

\leavevmode\hypertarget{ref-Dansereau2021SimJl}{}%
Dansereau, G., \& Poisot, T. (2021). SimpleSDMLayers.jl and GBIF.jl: A
Framework for Species Distribution Modeling in Julia. \emph{Journal of
Open Source Software}, \emph{6}(57), 2872.
\url{https://doi.org/10.21105/joss.02872}

\leavevmode\hypertarget{ref-Dominguez2020DecCon}{}%
Domínguez, L., \& Luoma, C. (2020). Decolonising Conservation Policy:
How Colonial Land and Conservation Ideologies Persist and Perpetuate
Indigenous Injustices at the Expense of the Environment. \emph{Land},
\emph{9}(3, 3), 65. \url{https://doi.org/10.3390/land9030065}

\leavevmode\hypertarget{ref-Dormann2010EvoCli}{}%
Dormann, C. F., Gruber, B., Winter, M., \& Herrmann, D. (2010).
Evolution of climate niches in European mammals? \emph{Biology Letters},
\emph{6}(2), 229--232. \url{https://doi.org/10.1098/rsbl.2009.0688}

\leavevmode\hypertarget{ref-Dunne2006NetStr}{}%
Dunne, J. A. (2006). The Network Structure of Food Webs. In J. A. Dunne
\& M. Pascual (Eds.), \emph{Ecological networks: Linking structure and
dynamics} (pp. 27--86). Oxford University Press.

\leavevmode\hypertarget{ref-Eichhorn2019SteDec}{}%
Eichhorn, M. P., Baker, K., \& Griffiths, M. (2019). Steps towards
decolonising biogeography. \emph{Frontiers of Biogeography},
\emph{12}(1), 1--7. \url{https://doi.org/10.21425/F5FBG44795}

\leavevmode\hypertarget{ref-Eklof2016PhyCom}{}%
Eklöf, A., \& Stouffer, D. B. (2016). The phylogenetic component of food
web structure and intervality. \emph{Theoretical Ecology}, \emph{9}(1),
107--115. \url{https://doi.org/10.1007/s12080-015-0273-9}

\leavevmode\hypertarget{ref-Fortin2021NetEco}{}%
Fortin, M.-J., Dale, M. R. T., \& Brimacombe, C. (2021). Network ecology
in dynamic landscapes. \emph{Proceedings of the Royal Society B:
Biological Sciences}, \emph{288}(1949), rspb.2020.1889, 20201889.
\url{https://doi.org/10.1098/rspb.2020.1889}

\leavevmode\hypertarget{ref-Galiana2021SpaSca}{}%
Galiana, N., Barros, C., Braga, J., Ficetola, G. F., Maiorano, L.,
Thuiller, W., Montoya, J. M., \& Lurgi, M. (2021). The spatial scaling
of food web structure across European biogeographical regions.
\emph{Ecography}, \emph{n/a}(n/a).
\url{https://doi.org/10.1111/ecog.05229}

\leavevmode\hypertarget{ref-Galiana2019GeoVar}{}%
Galiana, N., Hawkins, B. A., \& Montoya, J. M. (2019). The geographical
variation of network structure is scale dependent: Understanding the
biotic specialization of hostparasitoid networks. \emph{Ecography},
\emph{42}(6), 1175--1187. \url{https://doi.org/10.1111/ecog.03684}

\leavevmode\hypertarget{ref-Galiana2018SpaSca}{}%
Galiana, N., Lurgi, M., Claramunt-López, B., Fortin, M.-J., Leroux, S.,
Cazelles, K., Gravel, D., \& Montoya, J. M. (2018). The spatial scaling
of species interaction networks. \emph{Nature Ecology \& Evolution},
\emph{2}(5), 782--790. \url{https://doi.org/10.1038/s41559-018-0517-3}

\leavevmode\hypertarget{ref-Garland1999IntPhy}{}%
Garland, T., JR., Midford, P. E., \& Ives, A. R. (1999). An Introduction
to Phylogenetically Based Statistical Methods, with a New Method for
Confidence Intervals on Ancestral Values1. \emph{American Zoologist},
\emph{39}(2), 374--388. \url{https://doi.org/10.1093/icb/39.2.374}

\leavevmode\hypertarget{ref-Garlaschelli2018CovStr}{}%
Garlaschelli, D., Hollander, F. den, \& Roccaverde, A. (2018).
Covariance structure behind breaking of ensemble equivalence in random
graphs. \emph{Journal of Statistical Physics}, \emph{173}(3-4),
644--662. \url{https://doi.org/10.1007/s10955-018-2114-x}

\leavevmode\hypertarget{ref-GBIFSecretariat2021GbiBac}{}%
GBIF Secretariat. (2021). \emph{GBIF Backbone Taxonomy}.
\url{https://doi.org/10.15468/39omei}

\leavevmode\hypertarget{ref-Gerhold2015PhyPat}{}%
Gerhold, P., Cahill, J. F., Winter, M., Bartish, I. V., \& Prinzing, A.
(2015). Phylogenetic patterns are not proxies of community assembly
mechanisms (they are far better). \emph{Functional Ecology},
\emph{29}(5), 600--614. \url{https://doi.org/10.1111/1365-2435.12425}

\leavevmode\hypertarget{ref-Gravel2018BriElt}{}%
Gravel, D., Baiser, B., Dunne, J. A., Kopelke, J.-P., Martinez, N. D.,
Nyman, T., Poisot, T., Stouffer, D. B., Tylianakis, J. M., Wood, S. A.,
\& Roslin, T. (2018). Bringing Elton and Grinnell together: A
quantitative framework to represent the biogeography of ecological
interaction networks. \emph{Ecography}, \emph{0}(0).
\url{https://doi.org/10.1111/ecog.04006}

\leavevmode\hypertarget{ref-Grenie2021HarTax}{}%
Grenié, M., Berti, E., Carvajal-Quintero, J. D., Winter, M., \& Sagouis,
A. (2021). \emph{Harmonizing taxon names in biodiversity data: A review
of tools, databases, and best practices}.
\url{https://doi.org/10.32942/osf.io/e3qnz}

\leavevmode\hypertarget{ref-Grunig2020CroFor}{}%
Grünig, M., Mazzi, D., Calanca, P., Karger, D. N., \& Pellissier, L.
(2020). Crop and forest pest metawebs shift towards increased linkage
and suitability overlap under climate change. \emph{Communications
Biology}, \emph{3}(1, 1), 1--10.
\url{https://doi.org/10.1038/s42003-020-0962-9}

\leavevmode\hypertarget{ref-Halevy2009UnrEff}{}%
Halevy, A., Norvig, P., \& Pereira, F. (2009). The Unreasonable
Effectiveness of Data. \emph{IEEE Intelligent Systems}, \emph{24}(2),
8--12. \url{https://doi.org/10.1109/MIS.2009.36}

\leavevmode\hypertarget{ref-Halko2011FinStr}{}%
Halko, N., Martinsson, P. G., \& Tropp, J. A. (2011). Finding Structure
with Randomness: Probabilistic Algorithms for Constructing Approximate
Matrix Decompositions. \emph{SIAM Review}, \emph{53}(2), 217--288.
\url{https://doi.org/10.1137/090771806}

\leavevmode\hypertarget{ref-Herbert1965Dun}{}%
Herbert, F. (1965). \emph{Dune} (1st ed.). Chilton Book Company.

\leavevmode\hypertarget{ref-Hoffmann2019MacLea}{}%
Hoffmann, J., Bar-Sinai, Y., Lee, L. M., Andrejevic, J., Mishra, S.,
Rubinstein, S. M., \& Rycroft, C. H. (2019). Machine learning in a
data-limited regime: Augmenting experiments with synthetic data uncovers
order in crumpled sheets. \emph{Science Advances}, \emph{5}(4),
eaau6792. \url{https://doi.org/10.1126/sciadv.aau6792}

\leavevmode\hypertarget{ref-Holm2019DefBla}{}%
Holm, E. A. (2019). In defense of the black box. \emph{Science},
\emph{364}(6435), 26--27. \url{https://doi.org/10.1126/science.aax0162}

\leavevmode\hypertarget{ref-Hortal2015SevSho}{}%
Hortal, J., de Bello, F., Diniz-Filho, J. A. F., Lewinsohn, T. M., Lobo,
J. M., \& Ladle, R. J. (2015). Seven Shortfalls that Beset Large-Scale
Knowledge of Biodiversity. \emph{Annual Review of Ecology, Evolution,
and Systematics}, \emph{46}(1), 523--549.
\url{https://doi.org/10.1146/annurev-ecolsys-112414-054400}

\leavevmode\hypertarget{ref-Hutchinson2017CopSig}{}%
Hutchinson, M. C., Cagua, E. F., \& Stouffer, D. B. (2017).
Cophylogenetic signal is detectable in pollination interactions across
ecological scales. \emph{Ecology}, n/a--n/a.
\url{https://doi.org/10.1002/ecy.1955}

\leavevmode\hypertarget{ref-Jordano2016ChaEco}{}%
Jordano, P. (2016a). Chasing Ecological Interactions. \emph{PLOS Biol},
\emph{14}(9), e1002559.
\url{https://doi.org/10.1371/journal.pbio.1002559}

\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
Jordano, P. (2016b). Sampling networks of ecological interactions.
\emph{Functional Ecology}, \emph{30}(12), 1883--1893.
\url{https://doi.org/10.1111/1365-2435.12763}

\leavevmode\hypertarget{ref-Kawatsu2021AreNet}{}%
Kawatsu, K., Ushio, M., van Veen, F. J. F., \& Kondoh, M. (2021). Are
networks of trophic interactions sufficient for understanding the
dynamics of multi-trophic communities? Analysis of a tri-trophic insect
food-web time-series. \emph{Ecology Letters}, \emph{24}(3), 543--552.
\url{https://doi.org/10.1111/ele.13672}

\leavevmode\hypertarget{ref-Kefi2012MorMea}{}%
Kéfi, S., Berlow, E. L., Wieters, E. A., Navarrete, S. A., Petchey, O.
L., Wood, S. A., Boit, A., Joppa, L. N., Lafferty, K. D., Williams, R.
J., Martinez, N. D., Menge, B. A., Blanchette, C. A., Iles, A. C., \&
Brose, U. (2012). More than a meal\ldots{} integrating non-feeding
interactions into food webs: More than a meal \ldots. \emph{Ecology
Letters}, \emph{15}(4), 291--300.
\url{https://doi.org/10.1111/j.1461-0248.2011.01732.x}

\leavevmode\hypertarget{ref-Lamba2019DeeLea}{}%
Lamba, A., Cassey, P., Segaran, R. R., \& Koh, L. P. (2019). Deep
learning for environmental conservation. \emph{Current Biology},
\emph{29}(19), R977--R982.
\url{https://doi.org/10.1016/j.cub.2019.08.016}

\leavevmode\hypertarget{ref-Litsios2012EffPhy}{}%
Litsios, G., \& Salamin, N. (2012). Effects of Phylogenetic Signal on
Ancestral State Reconstruction. \emph{Systematic Biology}, \emph{61}(3),
533--538. \url{https://doi.org/10.1093/sysbio/syr124}

\leavevmode\hypertarget{ref-Machen2021ThiAlg}{}%
Machen, R., \& Nost, E. (2021). Thinking algorithmically: The making of
hegemonic knowledge in climate governance. \emph{Transactions of the
Institute of British Geographers}, \emph{46}(3), 555--569.
\url{https://doi.org/10.1111/tran.12441}

\leavevmode\hypertarget{ref-Maiorano2020Tet10}{}%
Maiorano, L., Montemaggiori, A., Ficetola, G. F., O'Connor, L., \&
Thuiller, W. (2020a). TETRA-EU 1.0: A species-level trophic metaweb of
European tetrapods. \emph{Global Ecology and Biogeography},
\emph{29}(9), 1452--1457. \url{https://doi.org/10.1111/geb.13138}

\leavevmode\hypertarget{ref-Maiorano2020DatTet}{}%
Maiorano, L., Montemaggiori, A., Ficetola, G. F., O'Connor, L., \&
Thuiller, W. (2020b). \emph{Data from: Tetra-EU 1.0: A species-level
trophic meta-web of European tetrapods} (Version 3, pp. 16596876 bytes)
{[}Data set{]}. Dryad. \url{https://doi.org/10.5061/DRYAD.JM63XSJ7B}

\leavevmode\hypertarget{ref-Marco2018ChaHum}{}%
Marco, M. D., Venter, O., Possingham, H. P., \& Watson, J. E. M. (2018).
Changes in human footprint drive changes in species extinction risk.
\emph{Nature Communications}, \emph{9}(1), 4621.
\url{https://doi.org/10.1038/s41467-018-07049-5}

\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
asymptotic network properties of spatial multi-trophic networks.
\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}

\leavevmode\hypertarget{ref-Mora2018IdeCom}{}%
Mora, B. B., Gravel, D., Gilarranz, L. J., Poisot, T., \& Stouffer, D.
B. (2018). Identifying a common backbone of interactions underlying food
webs from different ecosystems. \emph{Nature Communications},
\emph{9}(1), 2603. \url{https://doi.org/10.1038/s41467-018-05056-0}

\leavevmode\hypertarget{ref-Morales-Castilla2015InfBio}{}%
Morales-Castilla, I., Matias, M. G., Gravel, D., \& Araújo, M. B.
(2015). Inferring biotic interactions from proxies. \emph{Trends in
Ecology \& Evolution}, \emph{30}(6), 347--356.
\url{https://doi.org/10.1016/j.tree.2015.03.014}

\leavevmode\hypertarget{ref-MoseboFernandes2020MacLea}{}%
Mosebo Fernandes, A. C., Quintero Gonzalez, R., Lenihan-Clarke, M. A.,
Leslie Trotter, E. F., \& Jokar Arsanjani, J. (2020). Machine Learning
for Conservation Planning in a Changing Climate. \emph{Sustainability},
\emph{12}(18, 18), 7657. \url{https://doi.org/10.3390/su12187657}

\leavevmode\hypertarget{ref-Mouquet2012EcoAdv}{}%
Mouquet, N., Devictor, V., Meynard, C. N., Munoz, F., Bersier, L.-F.,
Chave, J., Couteron, P., Dalecky, A., Fontaine, C., Gravel, D., Hardy,
O. J., Jabot, F., Lavergne, S., Leibold, M., Mouillot, D., Münkemüller,
T., Pavoine, S., Prinzing, A., Rodrigues, A. S. L., \ldots{} Thuiller,
W. (2012). Ecophylogenetics: Advances and perspectives. \emph{Biological
Reviews}, \emph{87}(4), 769--785.
\url{https://doi.org/10.1111/j.1469-185X.2012.00224.x}

\leavevmode\hypertarget{ref-Nenzen2014Imp850}{}%
Nenzén, H. K., Montoya, D., \& Varela, S. (2014). The Impact of 850,000
Years of Climate Changes on the Structure and Dynamics of Mammal Food
Webs. \emph{PLOS ONE}, \emph{9}(9), e106651.
\url{https://doi.org/10.1371/journal.pone.0106651}

\leavevmode\hypertarget{ref-Neutel2002StaRea}{}%
Neutel, A.-M., Heesterbeek, J. A. P., \& de Ruiter, P. C. (2002).
Stability in Real Food Webs: Weak Links in Long Loops. \emph{Science},
\emph{296}(5570), 1120--1123.
\url{https://doi.org/10.1126/science.1068326}

\leavevmode\hypertarget{ref-Nokmaq2021AwaSle}{}%
No'kmaq, M., Marshall, A., Beazley, K. F., Hum, J., joudry, shalan,
Papadopoulos, A., Pictou, S., Rabesca, J., Young, L., \& Zurba, M.
(2021). {``Awakening the sleeping giant''}: Re-Indigenization principles
for transforming biodiversity conservation in Canada and beyond.
\emph{FACETS}, \emph{6}(1), 839--869.

\leavevmode\hypertarget{ref-Nost2021PolEco}{}%
Nost, E., \& Goldstein, J. E. (2021). A political ecology of data.
\emph{Environment and Planning E: Nature and Space}, 25148486211043503.
\url{https://doi.org/10.1177/25148486211043503}

\leavevmode\hypertarget{ref-OConnor2020UnvFoo}{}%
O'Connor, L. M. J., Pollock, L. J., Braga, J., Ficetola, G. F.,
Maiorano, L., Martinez-Almoyna, C., Montemaggiori, A., Ohlmann, M., \&
Thuiller, W. (2020). Unveiling the food webs of tetrapods across Europe
through the prism of the Eltonian niche. \emph{Journal of Biogeography},
\emph{47}(1), 181--192. \url{https://doi.org/10.1111/jbi.13773}

\leavevmode\hypertarget{ref-Pan2010SurTra}{}%
Pan, S. J., \& Yang, Q. (2010). A Survey on Transfer Learning.
\emph{IEEE Transactions on Knowledge and Data Engineering},
\emph{22}(10), 1345--1359. \url{https://doi.org/10.1109/TKDE.2009.191}

\leavevmode\hypertarget{ref-Park2004StaMec}{}%
Park, J., \& Newman, M. E. J. (2004). Statistical mechanics of networks.
\emph{Physical Review E}, \emph{70}(6), 066117.
\url{https://doi.org/10.1103/PhysRevE.70.066117}

\leavevmode\hypertarget{ref-Pedersen2017SigCol}{}%
Pedersen, E. J., Thompson, P. L., Ball, R. A., Fortin, M.-J., Gouhier,
T. C., Link, H., Moritz, C., Nenzen, H., Stanley, R. R. E., Taranu, Z.
E., Gonzalez, A., Guichard, F., \& Pepin, P. (2017). Signatures of the
collapse and incipient recovery of an overexploited marine ecosystem.
\emph{Royal Society Open Science}, \emph{4}(7), 170215.
\url{https://doi.org/10.1098/rsos.170215}

\leavevmode\hypertarget{ref-Perretti2013ModFor}{}%
Perretti, C. T., Munch, S. B., \& Sugihara, G. (2013). Model-free
forecasting outperforms the correct mechanistic model for simulated and
experimental data. \emph{Proceedings of the National Academy of
Sciences}, \emph{110}(13), 5253--5257.
\url{https://doi.org/10.1073/pnas.1216076110}

\leavevmode\hypertarget{ref-Pires2015PleMeg}{}%
Pires, M. M., Koch, P. L., Fariña, R. A., de Aguiar, M. A. M., dos Reis,
S. F., \& Guimarães, P. R. (2015). Pleistocene megafaunal interaction
networks became more vulnerable after human arrival. \emph{Proceedings
of the Royal Society B: Biological Sciences}, \emph{282}(1814),
20151367. \url{https://doi.org/10.1098/rspb.2015.1367}

\leavevmode\hypertarget{ref-Poelen2014GloBio}{}%
Poelen, J. H., Simons, J. D., \& Mungall, C. J. (2014). Global biotic
interactions: An open infrastructure to share and analyze
species-interaction datasets. \emph{Ecological Informatics}, \emph{24},
148--159. \url{https://doi.org/10.1016/j.ecoinf.2014.08.005}

\leavevmode\hypertarget{ref-Poisot2019EcoJl}{}%
Poisot, T., Belisle, Z., Hoebeke, L., Stock, M., \& Szefer, P. (2019).
EcologicalNetworks.jl - analysing ecological networks. \emph{Ecography}.
\url{https://doi.org/10.1111/ecog.04310}

\leavevmode\hypertarget{ref-Poisot2021GloKno}{}%
Poisot, T., Bergeron, G., Cazelles, K., Dallas, T., Gravel, D.,
MacDonald, A., Mercier, B., Violet, C., \& Vissault, S. (2021). Global
knowledge gaps in species interaction networks data. \emph{Journal of
Biogeography}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/jbi.14127}

\leavevmode\hypertarget{ref-Poisot2016StrPro}{}%
Poisot, T., Cirtwill, A. R., Cazelles, K., Gravel, D., Fortin, M.-J., \&
Stouffer, D. B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, \emph{7}(3), 303--312.
\url{https://doi.org/10.1111/2041-210X.12468}

\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
May 31). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}

\leavevmode\hypertarget{ref-Poisot2018IntRet}{}%
Poisot, T., \& Stouffer, D. B. (2018). Interactions retain the
co-phylogenetic matching that communities lost. \emph{Oikos},
\emph{127}(2), 230--238. \url{https://doi.org/10.1111/oik.03788}

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D. B., \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, \emph{124}(3), 243--251.
\url{https://doi.org/10.1111/oik.01719}

\leavevmode\hypertarget{ref-Price2003MacThe}{}%
Price, P. W. (2003). \emph{Macroevolutionary theory on macroecological
patterns}. Cambridge University Press.

\leavevmode\hypertarget{ref-Ray2021BioCri}{}%
Ray, J. C., Grimm, J., \& Olive, A. (2021). The biodiversity crisis in
Canada: Failures and challenges of federal and sub-national strategic
and legal frameworks. \emph{FACETS}, \emph{6}, 1044--1068.
\url{https://doi.org/10.1139/facets-2020-0075}

\leavevmode\hypertarget{ref-Reeve2016HowPar}{}%
Reeve, R., Leinster, T., Cobbold, C. A., Thompson, J., Brummitt, N.,
Mitchell, S. N., \& Matthews, L. (2016, December 8). \emph{How to
partition diversity}. \url{http://arxiv.org/abs/1404.6520}

\leavevmode\hypertarget{ref-Rosado2013GoiBac}{}%
Rosado, B. H. P., Dias, A., \& de Mattos, E. (2013). Going Back to
Basics: Importance of Ecophysiology when Choosing Functional Traits for
Studying Communities and Ecosystems. \emph{Natureza \&
Conservaç\textasciitilde ao Revista Brasileira de
Conservaç\textasciitilde ao Da Natureza}, \emph{11}, 15--22.
\url{https://doi.org/10.4322/natcon.2013.002}

\leavevmode\hypertarget{ref-Runghen2021ExpNod}{}%
Runghen, R., Stouffer, D. B., \& Dalla Riva, G. V. (2021).
\emph{Exploiting node metadata to predict interactions in large networks
using graph embedding and neural networks}.
\url{https://doi.org/10.1101/2021.06.10.447991}

\leavevmode\hypertarget{ref-Schoener1989FooWeb}{}%
Schoener, T. W. (1989). Food webs from the small to the large.
\emph{Ecology}, \emph{70}(6), 1559--1589.

\leavevmode\hypertarget{ref-Shlens2014TutPri}{}%
Shlens, J. (2014, April 3). \emph{A Tutorial on Principal Component
Analysis}. \url{http://arxiv.org/abs/1404.1100}

\leavevmode\hypertarget{ref-Solis-Lemus2017PhyPac}{}%
Solís-Lemus, C., Bastide, P., \& Ané, C. (2017). PhyloNetworks: A
Package for Phylogenetic Networks. \emph{Molecular Biology and
Evolution}, \emph{34}(12), 3292--3298.
\url{https://doi.org/10.1093/molbev/msx235}

\leavevmode\hypertarget{ref-Stock2021PaiLea}{}%
Stock, M. (2021). Pairwise learning for predicting pollination
interactions based on traits and phylogeny. \emph{Ecological Modelling},
14.

\leavevmode\hypertarget{ref-Stouffer2012EvoCon}{}%
Stouffer, D. B., Sales-Pardo, M., Sirer, M. I., \& Bascompte, J. (2012).
Evolutionary Conservation of Species' Roles in Food Webs.
\emph{Science}, \emph{335}(6075), 1489--1492.
\url{https://doi.org/10.1126/science.1216556}

\leavevmode\hypertarget{ref-Strong2014ImpNon}{}%
Strong, J. S., \& Leroux, S. J. (2014). Impact of Non-Native Terrestrial
Mammals on the Structure of the Terrestrial Mammal Food Web of
Newfoundland, Canada. \emph{PLOS ONE}, \emph{9}(8), e106264.
\url{https://doi.org/10.1371/journal.pone.0106264}

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., Forero-Muñoz, N. R., Higino, G., Mercier, B.,
Gonzalez, A., Gravel, D., Pollock, L., \& Poisot, T. (2021). A roadmap
towards predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \emph{376}(1837), 20210063.
\url{https://doi.org/10.1098/rstb.2021.0063}

\leavevmode\hypertarget{ref-Strydom2021SvdEnt}{}%
Strydom, T., Dalla Riva, G. V., \& Poisot, T. (2021). SVD Entropy
Reveals the High Complexity of Ecological Networks. \emph{Frontiers in
Ecology and Evolution}, \emph{9}.
\url{https://doi.org/10.3389/fevo.2021.623141}

\leavevmode\hypertarget{ref-Thessen2014KnoExt}{}%
Thessen, A. E., \& Parr, C. S. (2014). Knowledge extraction and semantic
annotation of text from the encyclopedia of life. \emph{PloS One},
\emph{9}(3), e89550.

\leavevmode\hypertarget{ref-Torrey2010TraLea}{}%
Torrey, L., \& Shavlik, J. (2010). Transfer learning. In \emph{Handbook
of research on machine learning applications and trends: Algorithms,
methods, and techniques} (pp. 242--264). IGI global.

\leavevmode\hypertarget{ref-Trojelsgaard2016EcoNet}{}%
Trøjelsgaard, K., \& Olesen, J. M. (2016). Ecological networks in
motion: Micro- and macroscopic variability across scales.
\emph{Functional Ecology}, \emph{30}(12), 1926--1935.
\url{https://doi.org/10.1111/1365-2435.12710}

\leavevmode\hypertarget{ref-Turak2017UsiEss}{}%
Turak, E., Brazill-Boast, J., Cooney, T., Drielsma, M., DelaCruz, J.,
Dunkerley, G., Fernandez, M., Ferrier, S., Gill, M., Jones, H., Koen,
T., Leys, J., McGeoch, M., Mihoub, J.-B., Scanes, P., Schmeller, D., \&
Williams, K. (2017). Using the essential biodiversity variables
framework to measure biodiversity change at national scale.
\emph{Biological Conservation}, \emph{213}, 264--271.
\url{https://doi.org/10.1016/j.biocon.2016.08.019}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
Upham, N. S., Esselstyn, J. A., \& Jetz, W. (2019). Inferring the mammal
tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology}, \emph{17}(12),
e3000494. \url{https://doi.org/10.1371/journal.pbio.3000494}

\leavevmode\hypertarget{ref-vanderHoorn2018SpaMax}{}%
van der Hoorn, P., Lippner, G., \& Krioukov, D. (2018). Sparse
Maximum-Entropy Random Graphs with a Given Power-Law Degree
Distribution. \emph{Journal of Statistical Physics}, \emph{173}(3-4),
806--844. \url{https://doi.org/10.1007/s10955-017-1887-7}

\leavevmode\hypertarget{ref-Vermaat2009MajDim}{}%
Vermaat, J. E., Dunne, J. A., \& Gilbert, A. J. (2009). Major dimensions
in food-web structure properties. \emph{Ecology}, \emph{90}(1),
278--282. \url{http://www.ncbi.nlm.nih.gov/pubmed/19294932}

\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
\url{https://doi.org/10.1002/ece3.1640}

\leavevmode\hypertarget{ref-Xing2021RisEco}{}%
Xing, S., \& Fayle, T. M. (2021). The rise of ecological network
meta-analyses: Problems and prospects. \emph{Global Ecology and
Conservation}, \emph{30}, e01805.
\url{https://doi.org/10.1016/j.gecco.2021.e01805}

\leavevmode\hypertarget{ref-Yeakel2014ColEco}{}%
Yeakel, J. D., Pires, M. M., Rudolf, L., Dominy, N. J., Koch, P. L.,
Guimarães, P. R., \& Gross, T. (2014). Collapse of an ecological network
in Ancient Egypt. \emph{PNAS}, \emph{111}(40), 14472--14477.
\url{https://doi.org/10.1073/pnas.1408471111}

\leavevmode\hypertarget{ref-Youden1950IndRat}{}%
Youden, W. J. (1950). Index for rating diagnostic tests. \emph{Cancer},
\emph{3}(1), 32--35.
\url{https://doi.org/10.1002/1097-0142(1950)3:1\%3C32::AID-CNCR2820030106\%3E3.0.CO;2-3}

\leavevmode\hypertarget{ref-Young2021BayInf}{}%
Young, J.-G., Cantwell, G. T., \& Newman, M. E. J. (2021). Bayesian
inference of network structure from unreliable data. \emph{Journal of
Complex Networks}, \emph{8}(6).
\url{https://doi.org/10.1093/comnet/cnaa046}

\leavevmode\hypertarget{ref-Young2007RanDot}{}%
Young, S. J., \& Scheinerman, E. R. (2007). Random Dot Product Graph
Models for Social Networks. In A. Bonato \& F. R. K. Chung (Eds.),
\emph{Algorithms and Models for the Web-Graph} (pp. 138--149). Springer.
\url{https://doi.org/10.1007/978-3-540-77004-6_11}

\leavevmode\hypertarget{ref-Zhu2006AutDim}{}%
Zhu, M., \& Ghodsi, A. (2006). Automatic dimensionality selection from
the scree plot via the use of profile likelihood. \emph{Computational
Statistics \& Data Analysis}, \emph{51}(2), 918--930.
\url{https://doi.org/10.1016/j.csda.2005.09.010}

\end{CSLReferences}

\end{document}